On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-12 appendix.tex
started summary

\section{Conversion formulae}
The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
to the total density $\rho$ and dust-fraction $\tepsilon$ by
\begin{align}
  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
    \epsilon},\\
  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
  \frac{\nabla^2\epsilon}{1+\epsilon} -
  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
\end{align}
and
\begin{align}
  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
 \nabla\tepsilon & =
 \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
 \nabla^2\tepsilon &=
 \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
 \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
\end{align}

>>>>

%\section{Conversion formulae}
%The gas density $\rhog$ and dust-to-gas ratio $\epsilon$ is related
%to the total density $\rho$ and dust-fraction $\tepsilon$ by
%\begin{align}
%  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \epsilon\right)},\\
%  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\epsilon}{1+
%    \epsilon},\\
%  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
%  \frac{\nabla^2\epsilon}{1+\epsilon} -
%  \frac{\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^2}, 
%\end{align}
%and
%\begin{align}
%  \tepsilon &= \frac{\epsilon}{1+ \epsilon},\\
% \nabla\tepsilon & =
% \frac{\nabla\epsilon}{\left(1+\epsilon\right)^2},\\
% \nabla^2\tepsilon &=
% \frac{\nabla^2\epsilon}{\left(1+\epsilon\right)^2} -
% \frac{2\left|\nabla\epsilon\right|^2}{\left(1+\epsilon\right)^3}. 
%\end{align}

\section{Dispersion relation for unstratified disks}
Unstratified disk with $\Phi = \Phi (r)$ so that $\p_zP = \p_z\rho =
0$. The background $\p_rP/\rho$ and $\tepsilon$ are constant. 
Fourier analyze in $r$ and $z$ so that $\p_z\to \ii k_z$ and
$\p_r\to \ii k_x$ when acting on perturbations. Denote $|\bm{k}|^2
\equiv k_x^2 + k_z^2$. Neglect variations in
background when compared with that of perturbations. Linearized
equations are, after eliminating the azimuthal velocity:

\begin{align}
  \ii \sigma W &=\nabla\cdot \bm{\dd v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \bm{\dd v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. 
\end{align}

Eliminate the velocity perturbations to obtain
\begin{align}
  \left(\ii \zeta \sigma - \tstop c_s^2 \tepsilon |\bm{k}|^2\right)Q &=
  \ii \left[
  \frac{P}{\rho}\sigma + k_x\tstop c_s^2 F_r\left(1 -
  2\tepsilon\right)\right]W. \\
   \sigma^2 \left( \kappa^2 - \sigma^2 - \ii k_x F_r\right)W &=
    \left(k_z^2\kappa^2 - \sigma^2|\bm{k}|^2\right)Q. 
\end{align}
Futher eliminate $Q$ and $W$ to obtain the dispersion relation
\begin{align}
  &\frac{\zeta}{c_s^2|\bm{k}|^2}\sigma^5 + \ii \tepsilon \tstop
  \sigma^4 - \left[ \frac{\zeta}{c_s^2|\bm{k}|^2} \left(\kappa^2 - \ii
  k_x F_r\right) + \left(1 - \tepsilon\right)\right]\sigma^3 - \ii
  \tstop \left[\tepsilon \kappa^2 - \ii k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 \notag\\ 
  &+ \left(1 -
  \tepsilon\right)\left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma +
  k_x \tstop F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align}
where the equation of state $P = c_s^2\left( 1 - \tepsilon\right)\rho$
was used. 
If we now set $\zeta = 0$ or consider $c_s^2\to \infty$ to obtain the
incompressible limit, and define $\tau_\mathrm{s} \equiv
\tstop/\left(1 - \tepsilon\right)$, we obtain: 
\begin{align*}
\ii \tepsilon \tstop
  \sigma^4 - \sigma^3 - 
  \tau_\mathrm{s} \left[ \ii \tepsilon \kappa^2 + k_x F_r\left( 1 -
  \tepsilon\right)\right]\sigma^2 
  + \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2\sigma + 
  k_x \tau_\mathrm{s} F_r\left(1 - 2\tepsilon\right)
  \left(\frac{k_z\kappa}{|\bm{k}|}\right)^2  = 0,
\end{align*}
as derived by \cite{jacquet11} and \cite{laibe14} for the streaming
instability.

### 2016-08-12 paper.tex

By re-casting the dusty problem into identical form to standard
hydrodynamics, we can effortlessly generalize previous stability
results for pure gas disks to the dusty case. We demonstrate this for
perfectly coupled dust in strictly and locally isothermal disks. We
show that sharp dust edges may destabilize protoplanetary disks.  
We also study the effect of dust-loading on the vertical shear
instability \citep[VSI,][]{nelson13,lin15,barker15}. We provide 
thermodynamical interpretations of our dusty stability calculations.

>>>>

By re-casting the dusty problem into identical form to standard
hydrodynamics, we can follow standard procedures to generalize previous stability 
results for pure gas disks to the dusty case. We demonstrate this for
perfectly coupled dust in strictly and locally isothermal disks. We
show that sharp dust edges may destabilize protoplanetary disks.  
We also study the effect of dust-loading on the vertical shear
instability \citep[VSI,][]{nelson13,lin15,barker15}. We provide 
thermodynamical interpretations of our dusty stability calculations.
